---
title: "MASH v FLASH analysis of GTEx data (random tests)"
output: 
  workflowr::wflow_html:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

I repeat the [previous experiment](MASHvFLASHgtex.html), this time using 20000 random tests from the GTEx dataset. An important difference from the "strong" dataset is that many more of these tests will be null.


## Results

I again pre-run the experiments and load the results from file.

```{r load_fits}
fl_diag <- readRDS("./output/MASHvFLASHrandom/fldiag.rds")
fl_t <- readRDS("./output/MASHvFLASHrandom/flt.rds")
m_diag <- readRDS("./output/MASHvFLASHrandom/mdiag.rds")
m_t <- readRDS("./output/MASHvFLASHrandom/mt.rds")
```

Comparisons of mean-squared error and confidence interval coverage are as follows. FLASH again outperforms MASH, but not quite so drastically it did on the "strong" tests.

```{r res1}
fl_fit_names <- c("pn.zero", "pn.OHF", "ash.zero", "ash.OHF")
method.names <- c(fl_fit_names, "mash")

all_t <- rbind(sapply(fl_t, unlist), unlist(m_t))
rownames(all_t) <- method.names

fl_mse <- sapply(fl_diag, function(x) {sapply(x, function(y) {y$mse})})
all_mse <- rbind(fl_mse, sapply(m_diag, function(x) {x$mse}))
rownames(all_mse) <- method.names

fl_ci <- sapply(fl_diag, function(x) {sapply(x, function(y) {y$ci})})
all_ci <- rbind(fl_ci, sapply(m_diag, function(x) {x$ci}))
rownames(all_ci) <- method.names

plot.order <- c(2, 4, 1, 3, 5)

boxplot(t(all_mse[plot.order, ]), ylim=c(0, 1), ylab = "MSE",
        main = "Mean-squared error")

boxplot(t(all_ci[plot.order, ]), ylim=c(0, 1), ylab = "coverage",
        main = "Proportion of 95% CIs containing true value")
```

The total time required by each method is as follows. 

```{r res2}
boxplot(t(all_t[plot.order, ]), ylim=c(0, 60), ylab="minutes",
        main="Time needed for fit and MSE and CI calculations")
```


## Conclusions

The "OHF" method again gives the best results, but not as convincingly as it did when fitting the "strong" dataset. And point-normal priors again give results that are only slightly worse than ash priors, but take about half of the time to fit.


## Code

The code is the same as in the [previous experiment](MASHvFLASHgtex.html#code) (to reproduce the above results,  uncomment the commented-out lines).
