---
title: 'MASH v FLASH GTEx analysis: nonnegative loadings'
output: 
  workflowr::wflow_html:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Here I fit MASH and FLASH objects to the "strong" GTEx dataset using the nonnegative loadings obtained in the [previous analysis](MASHvFLASHnn.html). See [below](#code) for code and [here](MASHvOHF.html#introduction_to_plots) for an introduction to the plots below.


## Fitting methods

The workflows are similar to those described [here](MASHvFLASHgtex2.html).

The workflow for FLASH is identical, except that I use the nonnegative loadings from the [previous analysis](MASHvFLASHnn.html) as my "data-driven" loadings. (To save time, I only use the loadings corresponding to "multi-tissue effects," since the other loadings are well represented by "canonical" loadings. Further, to aid interpretation, I normalize the loadings so that each has $\ell_\infty$-norm equal to 1.)

<!-- Additionally, I have manually adjusted some of the unique effects priors obtained by backfitting the loadings to the random dataset. As I point out [here](conclusions.html) (conclusion #3), some of these priors are effectively a point mass at zero, which causes an excessive amount of shrinkage in tests where a unique effect is visibly present. I adjust these priors so that they have sparsity ($\pi_0$) equal to $1/20000$ (since there are 20000 tests) and conditional variance ($1/a$) equal to one (this value was chosen somewhat arbitrarily, but is similar to estimated values of $1/a$ for other unique effects). (This adjustment is somewhat analogous to adding pseudocounts to count data.) -->

The workflow for MASH is also identical, except that I use rank-one matrices derived from the nonnegative loadings rather than rank-one matrices obtained using Extreme Deconvolution. More precisely, for each multi-tissue effect $\ell_i$, I add the matrix $\ell_i \ell_i^T$ to the list of matrices $U_i$ in the MASH mixture model.


## Results

I pre-run the code and load the fits from file.

```{r load_everything}
library(mashr)
devtools::load_all("/Users/willwerscheid/GitHub/flashr/")

source("./code/gtexanalysis.R")

gtex <- readRDS(gzcon(url("https://github.com/stephenslab/gtexresults/blob/master/data/MatrixEQTLSumStats.Portable.Z.rds?raw=TRUE")))
strong <- t(gtex$strong.z)

fpath <- "./output/MASHvFLASHgtex3/"
m_final <- readRDS(paste0(fpath, "m.rds"))
fl_final <- readRDS(paste0(fpath, "fl.rds"))

m_lfsr <- t(get_lfsr(m_final))
m_pm <- t(get_pm(m_final))

fl_lfsr <- readRDS(paste0(fpath, "fllfsr.rds"))
fl_pm <- flash_get_fitted_values(fl_final)
```

There is significant disagreement between the two fits:
```{r confusion}
signif_by_mash = as.vector(m_lfsr < 0.05)
signif_by_flash = as.vector(fl_lfsr < 0.05)

round(table(signif_by_mash, signif_by_flash) / length(signif_by_mash),
      digits = 2)
```


### Significant for FLASH, not MASH

As discussed in previous analyses (see, for example, [here](MASHvOHF.html#significant_for_ohf,_not_mash)), FLASH often finds a combination of a large unique effect and a small equal effect where MASH only finds the unique effect to be significant (or, more rarely, no effects).

```{r flash_not_mash}
# interesting.tests <- compare_methods(fl_lfsr, m_lfsr, fl_pm, m_pm)

par(mfrow=c(1, 2))

identical.plus.unique <- c(336, 7526, 2696, 611)
for (n in identical.plus.unique) {
  plot_test(n, fl_lfsr, fl_pm, "FLASH")
  plot_test(n, m_lfsr, m_pm, "MASH")
}
```


### Significant for MASH, not FLASH

In many cases, MASH finds all effects to be significant where FLASH only finds a single unique effect. What happens in these cases, I think, is that MASH puts nearly all posterior weight on a single data-driven covariance matrix. Since these covariance matrices are each rank-one with no entries that are exactly equal to zero, even tiny effects will appear significant (since they are sampled away from zero nearly all of the time).

In contrast, since the FLASH model allows for covariance structures that are linear combinations of these rank-one matrices, the FLASH posteriors are far less likely to follow these patterns of perfect correlation.

```{r mash_not_flash}
par(mfrow=c(1, 2))

mash.not.flash <- c(228, 65, 2903, 1460)
for (n in mash.not.flash) {
  plot_test(n, fl_lfsr, fl_pm, "FLASH")
  plot_test(n, m_lfsr, m_pm, "MASH")
}
```


### Different posterior means

As discussed [here](MASHvOHF.html#different_posterior_means), MASH often does poorly when effect sizes are uniformly large.

```{r diff_pms}
par(mfrow=c(1, 2))

diff_pms <- c(9956, 5293, 13168, 3680)
for (n in diff_pms) {
  plot_test(n, fl_lfsr, fl_pm, "FLASH")
  plot_test(n, m_lfsr, m_pm, "MASH")
}
```

## Code

Click "Code" to view the code used to obtain the fits used in this analysis.

```{r code, code=readLines("../code/MASHvFLASHgtex3.R"), eval=FALSE}
```

